A Partitioning Algorithm for Maximum Common (Connected) Subgraphs\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant numbers EP/K503058/1 and EP/M508056/1]}

### 2017-02-09 paper/paper.tex
More writing

\documentclass[letterpaper]{article}
\usepackage[pass,showframe]{geometry}

@@

\usepackage[ruled,vlined]{algorithm2e}
\usepackage{MnSymbol,wasysym}
\usepackage{latexsym}

@@

@@

\begin{abstract}
    We came up with an important new algorithm for the maximum common (connected)
    subgraph problem. This is an important problem, with at least two important
    applications.
\end{abstract}

@@

Given graphs $G=(\V(G), \E(G))$ and $H=(\V(H), \E(H))$, the \emph{Maximum Common
Subgraph (MCS)} problem is to find a graph with as many vertices as possible
that is isomorphic to induced subgraphs of both $G$ and $H$. Equivalently, MCS
is the problem of finding a maximum-cardinality set of pairs $M = \{(v_1, w_1),
\dots, (v_{|M|}, w_{|M|})\}$, with $v_i \in V(G)$ and $v_i \in V(H)$ for all $i$,
such that for each pair $i,j$ ($1 \leq i < j
\leq |M|$), $v_i$ and $v_j$ are adjacent in $G$ if and only if $w_i$ and $w_j$
are adjacent in $H$.

The most successful existing approaches to optimally solving MCS are
reformulation to the maximum clique problem and constraint programming
\cite{DBLP:conf/cp/McCreeshNPS16}. In this paper, we introduce a new branch and bound
algorithm for MCS which does not require a domain to be explicity maintained in
memory for each vertex in $\V(G)$. The algorithm maintains labellings of the
unmatched vertices in $\V(G)$ and $\V(H)$ such that a pair $(v, w)$ with $v \in \V(G)$
and $w \in \V(H)$ can be added to the matching if and only if $v$ and $w$ have the
same label.

@@

@@

@@

Suppose we are given the graphs $G$ and $H$ in Figure~\ref{fig:alg1}.  Graphs $G$ and $H$
have no common subgraph with five vertices. However, they do have common
subgraphs with four vertices; the mapping $\{1a, 2f, 3d, 5b\}$ is one example.

@@

\bigskip
\noindent
Phew!

\section{Heuristics}

We ran a small experiment comparing lots of heuristics such as \dots on random
instances. The one where you choose a colour class with as small a max(left size,
right size) as possible seemed to work best. ? Why?

@@

\section{Future Work}

We could create a hybrid of Lilybank and clique (for example, using Lilybank near the
top of the search tree, but switching to a clique encoding when fewer than some threshold
number of vertices remain to be selected). This might give us most of the benefits of
the clique encoding for labelled graphs, while avoiding the high memory cost and colouring
time of encoding the full instance.

@@

>>>>

\documentclass[letterpaper]{article}
%\usepackage[pass,showframe]{geometry}

@@

\usepackage[ruled,vlined]{algorithm2e}
\usepackage{MnSymbol,wasysym}
\usepackage{latexsym}
\usepackage{algpseudocode}

@@

\newcommand{\McSplit}{\textproc{McSplit}}

\newcommand{\nmax}{n_{\max}}

@@

\begin{abstract}
    Maximium Common Subgraph (MCS) is the problem of finding a
    graph of maximium size that is an induced subgraph of each of the two input
    graphs.  We introduce \McSplit, a branch and bound algorithm that
    uses a vertex-labelling scheme to compactly represent the subsets of
    vertices that may be matched together to create a larger common subgraph.
    The new algorithm explores the same search space as the existing CP-FC
    algorithm, but requires only \BigO{n^2} space in total and \BigO{n} time
    for each recursive call, greatly improving on all existing algorithms for
    the problem.

    Experimental results using two large graph databases show that \McSplit\ is
    more than an order of magnitude faster than the leading existing algorithms
    on instances involving undirected, unlabelled graphs, including the problem
    variant in which the subgraph must be connected.
\end{abstract}

@@

Given graphs $G=(\V(G), \E(G))$ and $H=(\V(H), \E(H))$, the \emph{Maximum Common
Subgraph (MCS)} problem is to find a graph with as many vertices as possible
that is isomorphic to induced subgraphs of both $G$ and $H$.

MCS has numerous applications, including finding the similarity between organic
molecules, \dots

@@

MCS is the problem of finding a maximum-cardinality set of pairs $M = \{(v_1,
w_1), \dots, (v_{|M|}, w_{|M|})\}$, with $v_i \in V(G)$ and $v_i \in V(H)$ for
all $i$, such that for each pair $i,j$ ($1 \leq i < j \leq |M|$), $v_i$ and
$v_j$ are adjacent in $G$ if and only if $w_i$ and $w_j$ are adjacent in $H$.

@@

\section{Prior Work}

The most successful existing approaches to solving MCS are reformulation to the
maximum clique problem on one hand, and constraint programming
\cite{DBLP:conf/cp/McCreeshNPS16} on the other. In the clique
\textit{association graph} encoding, a graph is created with one node for each
pair in the set $\{(u,v) \mid u \in \V(G), v \in \V(H) \}$ and edges representing
pairs of mappings that may be chosen together. A maximum clique in this graph
represents a maximum common subgraph of $G$ and $H$.

In the constraint programming approach, we have a variable corresponding to each
vertex in $G$, and each variable initially has a value in its domain for each vertex
in $H$ to which it may be mapped, plus an additional value $\bot$ signifying that
the vertex remains unmapped. A backtracking search is performed in which values
are assigned to variables, and the domains of unassigned variables are pruned.

In this paper, we introduce a
new branch and bound algorithm for MCS which does not require a domain to be
explicity maintained in memory for each vertex in $\V(G)$. The algorithm
maintains labellings of the unmatched vertices in $\V(G)$ and $\V(H)$ such that
a pair $(v, w)$ with $v \in \V(G)$ and $w \in \V(H)$ can be added to the
mapping if and only if $v$ and $w$ have the same label.

@@

We will use the graphs $G$ and $H$ in Figure~\ref{fig:alg1} to demonstrate the
algorithm.  These graphs have a maximum common subgraph with four vertices; the
mapping $\{1a, 2f, 3d, 5b\}$ is one example.

@@

\paragraph{Heuristics} We ran a small experiment comparing lots of heuristics
such as \dots on random instances. The one where you choose a colour class with
as small a max(left size, right size) as possible seemed to work best. ? Why?

@@

@@

We have introduced the \McSplit\ algorithm for Max Common Subgraph, and shown
that it explores the same search tree as CP-FC with greatly reduced effort at
each search node.  The new algorithm is more than an order of
magnitude faster than the previous state of the art for unlabelled, undirected
instances.

\paragraph{Future work} We could create a hybrid of Lilybank and clique (for
example, using Lilybank near the top of the search tree, but switching to a
clique encoding when fewer than some threshold number of vertices remain to be
selected). This might give us most of the benefits of the clique encoding for
labelled graphs, while avoiding the high memory cost and colouring time of
encoding the full instance.
